Language Bias in a Multilingual Tweet Corpus

### 2018-05-17 paper.tex
Updates from ShareLaTeX

Our base assumption is that geo-located tweets is a representative and unbiased sample of Latvian Twitter. Thus, an extended collection should exhibit similar properties as the geo-located one. The main property we are interested in this work is the language proportion. We hypothesize, that an unbiased extended collection should have similar language distribution as the initial.

To study the collection in more details, we defined 9 topics and manually search for relevant tweets.

@@

The analysis of the collection shows that it is biased toward content in Latvian: proportionally there are more tweets in Latvian than in Russian and English. Also, Latvian is more dominant language than expected in all but one topic, the exception being the topic about New Year celebration.

@@

Building an unbiased collection is difficult.  In our case the main reason of introduced bias, might be that Russian and English content is less news-oriented and much more informal. In other words, while it's common to discuss current news in Latvian by engaging with media, the tweets in Russian and English might be more personal and friend-oriented.

@@

Over the period from 15 April 2017 to \hl{8 May 2018} the initial set of \hl{1\,708\,236} tweets was collected from the \texttt{POST status/filter} endpoint of the Twitter Streaming API.%\footnotemark{}
%
%\footnotetext{\url{https://developer.twitter.com/en/docs/tweets/filter-realtime/api-reference/post-statuses-filter}}
%
The \texttt{locations} parameter was set to the bounding box of Riga, the capital of Latvia.%\footnotemark{}
%
%\footnotetext{The coordinates of the bounding box are \texttt{23.9325829, 56.8570671, 24.3247299, 57.0859184}.}
%
In addition to the location, 420 accounts were tracked.%\footnotemark{} %\hl{The whole list of screen names together with user ID is available in the supplement.}
%\footnotetext{In this paper, by ``tracking users'', we mean that their user IDs where passed to the Twitter API to collect tweets that they created, or the tweets that are replies by other users to their tweets. Refer to \url{https://developer.twitter.com/en/docs/tweets/filter-realtime/guides/basic-stream-parameters\#follow} for a complete description. Keep in mind that this is different from the case when a user follows another user.}
%
The accounts are mostly Latvian (that is, coming from Latvia, but not necessary produce content in Latvian) news outlets, politicians, businesses, artists and sport clubs.

>>>>

Our base assumption is that geo-located tweets are a representative and unbiased sample of Latvian Twitter. Thus, an extended collection should exhibit similar properties as the geo-located one. The main property we are interested in this work is the language proportion. We hypothesize, that an unbiased extended collection should have similar language distribution as the initial.

To study the collection in more detail, we defined 9 topics and manually search for relevant tweets.

@@

The analysis of the collection shows that it is biased toward content in Latvian: proportionally there are more tweets in Latvian than in Russian and English. Also, Latvian is more dominant language than expected in all but one topic, the exception being the topic about New Year celebration.
# IMS I feel like this paragraph is hard to understand because the notion of "topics" hasn't been really defined yet.  Since this is the introduction, the solution might be to make this statement more general.

@@

Building an unbiased collection is difficult.  In our case the main reason of introduced bias, might be that Russian and English content is less news-oriented and much more informal. In other words, while it's common to discuss current news in Latvian by engaging with media, the tweets in Russian and English might be more personal and friend-oriented.
# IMS is this a hypothesis you intend to test here?  Or is this future work?  Or what?

@@

Over the period from 15 April 2017 to \hl{8 May 2018} the initial set of \hl{1\,708\,236} tweets was collected from the \texttt{POST status/filter} endpoint of the Twitter Streaming API.%\footnotemark{}
%
%\footnotetext{\url{https://developer.twitter.com/en/docs/tweets/filter-realtime/api-reference/post-statuses-filter}}
%
The \texttt{locations} parameter was set to the bounding box of Riga, the capital of Latvia.%\footnotemark{}
%
%\footnotetext{The coordinates of the bounding box are \texttt{23.9325829, 56.8570671, 24.3247299, 57.0859184}.}
% IMS according to http://worldpopulationreview.com/countries/latvia-population/, the population of Riga is less than half the total population of Latvia.  However, Twitter only allows bounding boxes for location to be so big, so you can't bound the whole country.

In addition to the location, 420 accounts were tracked.%\footnotemark{} %\hl{The whole list of screen names together with user ID is available in the supplement.}
%\footnotetext{In this paper, by ``tracking users'', we mean that their user IDs where passed to the Twitter API to collect tweets that they created, or the tweets that are replies by other users to their tweets. Refer to \url{https://developer.twitter.com/en/docs/tweets/filter-realtime/guides/basic-stream-parameters\#follow} for a complete description. Keep in mind that this is different from the case when a user follows another user.}
%
The accounts are mostly Latvian (that is, coming from Latvia, but not necessary produce content in Latvian) news outlets, politicians, businesses, artists and sport clubs.
